---
title: 'Statistical report: Association of Beta-blocker Use and Heart Failure Hospitalization and Cardiovascular Disease Mortality in Patients with HFpEF >=50%'
author: 'Statistician: Lina Benson'
  
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    number_sections: yes
link-citations: yes
bibliography: references.bib
nocite: '@*'
---

\newpage 
\tableofcontents 
\listoffigures
\listoftables
\newpage


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, include = TRUE, comment = "",
  warning = FALSE, message = FALSE, fig.pos = "H",
  fig.path = "../output/figs/"
)
options(knitr.kable.NA = "")
```

```{r adjust_directory_if_needed, include=FALSE}
# Uncomment lines below if rmd file is placed in a subdirectory
knitr::opts_knit$set(root.dir = normalizePath("../"))
```

```{r load_project}
# 1. Set options in config/global.dcf
# 2. Load packages listed in config/global.dcf
# 3. Import functions and code in lib directory

ProjectTemplate::reload.project()

cacheon <- TRUE
```             

# Data handling

## Data source

SHFDB3, https://kiheartfailure.github.io/shfdb3/, v 3.1.4. 

## Inclusion/exclusion criteria

```{r flow}
mykable(flow, caption = "Flowchart")
```

First patient in: `r min(pdata$shf_indexdtm)` and last patient in: `r max(pdata$shf_indexdtm)`. 
Median age (IQR): `r pdata %>% summarise(med = dF(median(shf_age), dig = 1),
                                             q1 = dF(quantile(shf_age, probs = 0.25), dig = 1),
                                             q3 = dF(quantile(shf_age, probs = 0.75), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", q1, "-", q3, ")")) %>%
                                   pull(out)` and 
`r pdata %>% count(shf_sex) %>%
  mutate(perc = dF(n / sum(n) * 100, 1)) %>%
  filter(shf_sex == "Female") %>%
  pull(perc)`% females.                                   

# Statistical analysis 

## General

All analyses were performed using `r sessionInfo()$R.version$version.string` [@r]. 
The level of significance is set to 5%, two-sided. No adjustment for multiple 
comparisons were made and therefore the results should be viewed with care.

## Missing data

Missing data was imputed with multiple imputation using mice [@mice] for 10 
datasets and 10 iterations. Variables included in the model are indicated in 
Table \ref{tab:tab1}. The primary outcome, first heart failure hospitalization,
was included as the Nelson-Aalen estimator. Bbl use is not included in the imputation model.  

## Propensity scores

A propensity score for treatment with bbl was estimated for each patient with 
logistic regression for each of the 10 imputed datasets using with variables 
indicated in Table \ref{tab:tab1}. Continuous variables were modeled using 
cubic splines with 3 df. 2:1 matching without replacement [@match] 
was thereafter performed on the average of the resulting 10 ps [@psmatch]. Matching was 
allowed if the propensity score differed by 0.01 or less. The ability of the ps 
matching to balance the baseline characteristics was assessed by 
standardized mean differences.  

## Baseline characteristics

```{r, child = "../src/tab1.Rmd"}

```

```{r, child = "../src/tabtreat.Rmd"}

```

## Association between bbl and outcomes 

The following outcomes are considered:

- First HF hospitalization (primary)
- Cardiovascular death (CVD)
- First all-cause hospitalization
- All-cause mortality

The main analysis is the ps matched analysis including NT-proBNP with bbl 
according to the ITT principle. All analyses are on this population unless 
stated otherwise. 

Data were censored at 2018-12-31 or death/emigration, 
so this assumes that the risk of cv mortality and hf hospitalization 
would have been the same for a censored patient had the censoring not occurred.

Incidence per 1000 py was calculated with 95% Poisson confidence intervals. 

The primary and secondary outcomes were presented with the 1 - Kaplan-Meier curves 
and tested with the log-rank test for the matched population 
(note that the log-rank test is theoretically 
redundant since the cox regression gives the same information). 

Cox proportional hazards regressions were used to evaluate the association 
between bbl and the respective outcomes. 
Continuous variables were modeled using cubic splines with 3 df.
In the matched cohort the matched pairs were modeled using a frailty term. 

The mini adjusted models includes age and sex. The partial adjusted models 
includes `r paste(modvarspartial, collapse = ", ")`. The full adjusted models 
includes the variables indicated in Table \ref{tab:tab1}. 

### Consistency analysis

#### Crossover

In the main analysis bbl is considered according to the ITT principle 
(bbl at baseline). As a consistency analysis patients were censored at crossover. 
In untreated patients, crossover was defined as a new bbl (ATC C07) dispension 
in the Dispensed Drug Registry and for treated patients, crossover was defined 
as no bbl dispension within 6 months from baseline or from last dispension, 
where crossover was set 3 months from index or last refill. 

#### Excluding NT-proBNP

Since the % missing observations with NT-proBNP was high, analyses were also 
performed excluding NT-proBNP in the imputation, matching and outcomes analyses 
resulting in N = `r matchp_sens %>% filter(shf_bbl == "No") %>% count()` 
matched patients without bbl.  

#### Competing risk

The analyses were also performed using a sub-distributional hazards model [@fg] 
were death was treated as a competing event. 

#### EF subgroups

The analyses were also performed including only EF >= 55% 
N =
`r paste(pdata %>% filter(shf_efproc >= 55) %>% count(shf_bbl) %>% pull(n), collapse = "/")` 
(bbl no/yes respectively) and 60% N = 
`r paste(pdata %>% filter(shf_efproc >= 60) %>% count(shf_bbl) %>% pull(n), collapse = "/")`
for the crude, mini and partially adjusted models. The fully adjusted models, 
including the propensity score, were 
not performed since there were not enough observations to support the models. 
Since continuous EF has been collected since approximately April 2017 the population
is highly selected with a shorter follow-up and therefore also an 
analysis on EF >= 50% with continuous EF was performed N = 
`r paste(pdata %>% filter(shf_efproc >= 50) %>% count(shf_bbl) %>% pull(n), collapse = "/")`.

#### Other subgroups

Subgroups for selected other variables was depicted in a Forest plot for the matched 
cohort using interaction effects. 

### Assumptions

The proportional hazards assumption was investigated using the scaled Schoenfeld 
residuals (cox.zph in [@survival-package]) for the primary and secondary outcome 
(bbl according to ITT). Possible outliers were visually inspected by plotting the dfbetas. 
=> shf_location, shf_rasarni
were deemed not to have proportional hazards and therefore stratified for in the cox regressions.

```{r, child = "../src/km.Rmd"}

```

```{r, child = "../src/outtab.Rmd"}

```

```{r, child = "../src/outtabsub.Rmd"}

```

\clearpage
\newpage

## NT-proBNP

The plot is performed on the whole cohort (not the matched cohort). 

```{r, child = "../src/boxplot.Rmd"}

```

\clearpage
\newpage

# Reproducibility

## R session information {#sessioninfo}

`r sessionInfo()$R.version$version.string` [@r] is used, for more details see below.

```{r sessinfo}
sessionInfo()
```

## R code

The R code for all data handling and statistical analyses are found: 
https://github.com/KIHeartFailure/bblEFover50. On publication the repository 
will be made public so it is possible to 
link to it from the resulting article for increased transparency and code sharing.
No data or output is stored in the repository. 

# References